---
title: "Venice deaths on Maps"
author: "G. Lazzari and G. Colavizza"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

```{r}
# clean env. vars
rm(list=ls()) 

# packages for maps
library(maptools) # gpclibPermit()
library(tmap)
library(rgdal)
library(oce)
library("RColorBrewer")

# basic packages
library("ggplot2")
library(reshape2)
library(dplyr)
library(lubridate)
```

# Load shap file for map skeleton
```{r}
shap = readOGR('../DataGIS/1729_PAR_CONF.shp')

# 'id' does not seem to be useful --> remove it
shap@data <- shap@data['NAME']
```

# Original deaths on map

## Load, clean  dataset

```{r}
original_deaths <- read.csv('../dataset/ALL_utf.csv')
# mind that NA have to be set to 0, ...
# otherwise the aggregrated sum will be reduced to NA!, but...
original_deaths[is.na(original_deaths)] <- 0 # cleaning NAs >> we'll set them back later to NA (***)

# fix dates
original_deaths$Date <- as.POSIXct(original_deaths$Date,format = '%m/%d/%Y')
my_shift <- as.POSIXct('1629-01-01')
original_deaths$Date <- original_deaths$Date - difftime(original_deaths$Date[1],my_shift )
```


## aggregate over months

(and set back NAs)
```{r}
# melt parishes
melted_df  =  melt(original_deaths,'Date',variable.name = 'parish',value.name = 'deaths')

# group over months, sum ...
# credits to: https://ro-che.info/articles/2017-02-22-group_by_month_r 
agg_melt_df = melted_df %>% group_by(month=floor_date(Date, "month"),parish=parish) %>% 
      summarise(deaths = sum(deaths))

# ...and un-melt again
agg_df = as.data.frame(t(dcast(agg_melt_df,month~parish,drop = FALSE)))

#...and remove useless row
agg_df = agg_df[-1,]

# save parishes names as they will be canceledby next step
parish_names = rownames(agg_df)

# trasform all data to numeric 
agg_df = as.data.frame(sapply(agg_df,function(x){as.numeric(as.character(x))}))

#  (***) set parish with no deaths on the whole t-w to NAs!
agg_df[rowSums(agg_df) == 0,] = NA

# create name col
agg_df$NAME = parish_names

# remove vars of intermediate steps
remove(melted_df,agg_melt_df,parish_names)
```


## sum over whole time window 
(don't run this section if you are not interested)

create cols for sums and names and drop the rest (and set back NAs)

```{r}
original_summed <- as.data.frame(colSums(original_deaths[,2:ncol(original_deaths)]))
colnames(original_summed) <- 'Total_deaths'
original_summed$NAME <- rownames(original_summed)
remove(original_deaths)

# remember we had to set NAs to 0 to avoid missing some data). Indeed now we have lost all NAs:
sum(is.na(original_summed))
# so we set it back to NA:
original_summed[original_summed==0] = NA
sum(is.na(original_summed))
```

## merge deaths data into shape object

fix col names to match names in shapes: 
```{r} 
#replace spaces with dots 
shap@data$NAME <- sapply(shap@data$NAME, function(x) gsub('\\s','.',x))
```

fix col names to match names in dataset summed  >> over all time-window <<
```{r}
# the only problematic one is S..Giacomo.dell.Orio --> fixed it: 
original_summed$NAME[original_summed$NAME=='S..Giacomo.dell.Orio'] <- 'S..Giacomo.dell\'Orio'

# check (in)compatibility of names 
sum(!(original_summed$NAME %in% shap@data$NAME)) # -- not bad ;), 0 mismatch!

### view mismatches ### 
View(shap@data[!(shap@data$NAME %in% original_summed$NAME),]) # 15 parishes 
View(original_summed[!(original_summed$NAME %in% shap@data$NAME),])
```


fix col names to match names in dataset aggregated >> over month <<
```{r}
# the only problematic one is S..Giacomo.dell.Orio --> fixed it: 
agg_df$NAME[agg_df$NAME=='S..Giacomo.dell.Orio'] <- 'S..Giacomo.dell\'Orio'
## check (in)compatibility of names 
sum(!(agg_df$NAME %in% shap@data$NAME)) # -- not bad ;), 0 mismatch!

### view mismatches ### 
View(shap@data[!(shap@data$NAME %in% agg_df$NAME),]) # 15 parishes have no deaths counts -- it's fine
View(agg_df[!(agg_df$NAME %in% shap@data$NAME),1:2]) 
```

merge data summed over --> all time window <--
```{r}
# keep track of original sorting -- before merging!!
shap@data$id = rownames(shap@data)

# actual merging
shap@data = merge(shap@data,original_summed,
                  by = 'NAME',all.x = TRUE,all.y = TRUE,sort = FALSE)

# set id as numeric 
shap@data$id = as.numeric(as.character(shap@data$id))
# order df using previous id 
shap@data = shap@data[order(shap@data$id),]
```

merge data aggregated --> over months  <--
```{r}

# keep track of original sorting -- before merging!!
shap@data$id = rownames(shap@data)

# actual merging
shap@data = merge(shap@data,agg_df,
                  by = 'NAME',all.x = TRUE,sort = FALSE)

# set id as numeric 
shap@data$id = as.numeric(as.character(shap@data$id))
# order df using previous id 
shap@data = shap@data[order(shap@data$id),]

# set deaths data as numeric
shap@data[,3:ncol(shap@data)] = sapply(shap@data[,3:ncol(shap@data)],as.numeric)

# fix col names
months_names <- c(
    'january 1629','february 1629','march 1629','april 1629','may 1629','june 1629','july 1629','august 1629','september 1629','october 1629','november 1629','december 1629',    
    'january 1630','february 1630','march 1630','april 1630','may 1630','june 1630','july 1630','august 1630','september 1630','october 1630','november 1630','december 1630',
    'january 1631','february 1631','march 1631','april 1631','may 1631','june 1631','july 1631','august 1631','september 1631','october 1631','november 1631','december 1631'
)

colnames(shap@data) <- c(c('NAME','id'),months_names)


View(shap@data)
```

## Plot cases on map 
(doc for space analysis/plot in R: https://cran.r-project.org/doc/contrib/intro-spatial-rl.pdf )


in order to simplify the plot, we keep only labels for parishes with: N deaths > cut-off 
```{r}
# names_cutoff = 

my_name = function(x,cutoff=1000){
    
    if ((shap@data$Total_deaths[shap@data$NAME==x] > cutoff) & 
        !is.na(shap@data$Total_deaths[shap@data$NAME==x])){
        return(x)
    }
    else{
        return(NA)
    }
    
}

shap@data$names_cutoff = sapply(shap@data$NAME,my_name)
```



map cases >>> by month <<<
```{r}
pdf('map_monthly_cases.pdf',height = 6, width = 12)

# (parishes names makes plot unreadable)
qtm(shap ,fill = colnames(shap@data)[20:ncol(shap@data)],ncol = 5)

dev.off()

### testing on single month 
# qtm(shap, text = 'NAME', text.size = 0.5, fill = c('january 1631'))
```

map with labels >>> total cases  <<< 
```{r}
pdf('map_total_cases_with_labels.pdf',height = 6, width = 12)

tm_shape(shap) + tm_fill('Total_deaths') + tm_text('NAME', size = 0.5 , col = 'black') + tm_layout(bg.color="lightblue") 

## saving
dev.off()
```

here we fix the text going over boundary -- in the two-row name
(fttb, we change only the only problematic one)
```{r}
shap@data$names_cutoff[shap@data$NAME=='S..Pietro.di.Castello'] = 'S..Pietro\ndi.Castello'

View(shap@data[shap@data$NAME=='S..Pietro.di.Castello',])
```

map with labels >>> total cases  <<<  -- version with less (cutoff) labels 
```{r}
pdf('map_total_cases_with_labels_cutoff.pdf',height = 6, width = 12)
# qtm(shap,'Total_deaths',text = 'names_cutoff', text.size = 0.8,text.col = 'black')

tm_shape(shap) + tm_fill('Total_deaths') + tm_text('names_cutoff', size = 0.8 , col = 'black') + tm_layout(bg.color="lightblue") 


## saving
dev.off()
```


map total cases without labels
```{r}
# pdf('map_total_cases_no_labels.pdf',height = 6, width = 12)
qtm(shap,'Total_deaths')

## saving
dev.off()
```
